---
title: "Homework 3"
author: "Chan Pok Wah"
output:
  html_document:
    highlight: tango
    theme: paper
    toc: yes
    toc_depth: 3
---

Before proceeding to solve the problems, make sure the `tidyverse` collection is installed in your computer. 

Note that some of the code blocks in this assignment contain the expression `eval = FALSE`. This tells R Markdown to display the code contained in the block, but not to evaluate it. Remove it if you want to turn off this behavior.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
```

### Problem 1: Plotting summary statistics with histograms

The following code generates three random vectors and then produce their histograms, respectively:

```{r}
# sets seed to make sure everyone's random vectors are generated the same
set.seed(201611) 
list.random <- list(x = rnorm(50), y = runif(65), z = rt(100, df = 1.5)) 

# here's a figure showing histograms of the data
par(mfrow = c(1,3))
hist(list.random$x, col = 'lightgrey') 
hist(list.random$y, col = 'salmon') 
hist(list.random$z, col = 'steelblue')
```

##### (a) Denoting the means of the three vectors in the respective histograms with vertical lines.


```{r}
# here's a figure showing histograms of the data
par(mfrow = c(1,3))

# edit here
hist(list.random$x, col = 'lightgrey') 
abline(v=mean(list.random$x), col="black", lwd=2)
hist(list.random$y, col = 'salmon')
abline(v=mean(list.random$y), col="black", lwd=2)
hist(list.random$z, col = 'steelblue')
abline(v=mean(list.random$z), col="black", lwd=2)
```
 

##### (b) Adding density curves to approximate the three histograms.

*Tip: use a probability scale for histograms to allow the density curve to be added.*

```{r}
# here's a figure showing histograms of the data
par(mfrow = c(1,3))
# edit here
hist(list.random$x, prob=TRUE, col="lightgrey")# prob=TRUE for probabilities not counts
lines(density(list.random$x), col="blue", lwd=2) # add a density estimate with defaults
hist(list.random$y, prob=TRUE, col="salmon")# prob=TRUE for probabilities not counts
lines(density(list.random$y), col="blue", lwd=2) # add a density estimate with defaults
hist(list.random$z, prob=TRUE, col="steelblue")# prob=TRUE for probabilities not counts
lines(density(list.random$z), col="blue", lwd=2) # add a density estimate with defaults
```




## Problem 2: Plotting financial time series for stock prices

`weeklyprice.csv` contains time series data for stock prices for three large Internet companies (Facebook, Google, and Amazon) from Oct. 30, 2012 to Oct. 30, 2017. Load the dataset to create a tibble named `weekly` using `read_csv()` from the `readr` package. 



```{r}
weekly <- read_csv("https://drive.google.com/uc?export=download&id=1uAYO3mJYsRPAPXn-ZE5llEYebnZhcdqJ", col_names = TRUE, col_types = cols())
weekly %>% print(width = Inf)
```


It has 3 rows and 263 columns. Each row stores weekly closing prices for a firm's stock during the aforementioned period. You may notice that the data has the wide format that is perhaps not suitable for plotting.

##### (a) Tidying the data

Tranform the data from the wide format to the long format. The resulting tibble, which is called `weekly_long`, should have three columns, one called `company`, one called `date`, and one called `price`.

```{r}
# edit here
weekly_long <- gather(weekly, date, price, "2012-10-29":"2017-10-30")
print(weekly_long)

```



##### (b) Plotting multiple data series in one plot

Plot time series of stock prices for the three companies with seperate curves, each having a different color. Try to reproduce the plot given in the assignment description.

*Tip: use the following code to customize the axis representing calendar dates: `axis.Date(side=1, at=seq(min(weekly_long$date), max(weekly_long$date), by="6 mon"), format="%m-%Y",cex.axis=0.7)`*

```{r}
FB <- weekly_long[weekly_long$company == "FB",]
GOOG <- weekly_long[weekly_long$company == "GOOG",]
AMZN <- weekly_long[weekly_long$company == "AMZN", ]
plot(as.Date(FB$date), FB$price, type="l", xaxt="n", xlab="Date",ylab="Closing Price", main="Time Series of Stock Prices",cex.axis=0.7, xlim=c(as.Date(min(weekly_long$date)), as.Date(max(weekly_long$date))), 
  ylim=c(min(weekly_long$price), max(weekly_long$price)), col="red")
lines(as.Date(GOOG$date), GOOG$price, col="blue", type="l")
lines(as.Date(AMZN$date), AMZN$price, col="green", type="l") 
minDate <- min(weekly_long$date)
maxDate <- max(weekly_long$date)
axis.Date(side=1, at=seq(from = as.Date(minDate), as.Date(maxDate), by="6 mon"), format="%m-%Y",cex.axis=0.7)
legend("topleft", legend = c("FB", "GOOG", "AMZN"), lty = 1, col=c("red", "blue", "green"))

```
 

## Problem 3: Practicing the `dplyr` package with flights data

Download `flights.csv` and `airports` from the Web to create two tibbles named `flights` and `airports`: 

```{r}
flights<-read_csv("https://drive.google.com/uc?export=download&id=1zAE0s3FA0dvhahex5J44jT5azutDacm0", col_names = TRUE, col_types = cols())
airports<-read_csv("https://drive.google.com/uc?export=download&id=1LLthXwTC-wqXnsMPU9-nvknIAMixkmH3", col_names = TRUE, col_types = cols())
## glimpse() is a transposed version of print
glimpse(flights)
summary(flights)
```


The `flights` dataset is about all the flights that departed from New Yord City (i.e. airports JFK, LGA or EWR) in 2013. In particular, the interest lies in the following variables:

- `hour`, `minute`: the hour and minute of the departure
- `arr_delay`: the arrival delay of the incoming plane (in minutes)
- `dest`: the destination

Note that several variables have ***missing values***.

##### (a) Creating new variables

Create a new variable which encodes a given `hour` and `minute` as one decimal number, i.e. time in hours; that is, for example, 2 hours 45 minutes should be coded to 2.75 hours, because 45 minutes is 45 minutes * (1 hour / 60 minutes) = 0.75 hours. Name this new variable `time` in `flights`.

```{r}
# edit here
flights$time <- flights$hour + flights$minute / 60.0
glimpse(flights)

```


##### (b) Plotting the destinations.

Calculate the average value of the arrival delay (`arr_delay`) and the number of departing flights (`n`) for each destination (`dest`) and name the resulting data frame `delay.per.dest`.

Here you can use the summary function `n()` to find the count for each group. When calculating the mean, you should be aware of the existence of the missing values. 

```{r}
# edit here

delay.per.dest <- summarise(group_by(flights, dest),
                   arr_delay = mean(flights$arr_delay, na.rm=TRUE),
                   n = n())
delay.per.dest
```

The `airports` dataset contains the coordinates (`lon`, `lat`) of the 1,458 airports. `faa` stands for FAA airport code.

```{r}
airports
```

Merge the tibble `delay.per.dest` and `airports` in order to add the coordinates (`lon`, `lat`)  of the airports to `delay.per.dest` using `left_join()` in `dplyr` (type `?left_join` to see how it works).

```{r}
# edit here
delay.per.dest <- left_join(delay.per.dest, airports, by = c("dest" = "faa"))

delay.per.dest
```

Create a scatterplot of the latitude against the longitude and scale the points according to the number of departing planes.

```{r eval=T}
plot(lat ~ lon, data = delay.per.dest, pch = 19, cex = n / 6000)
```

<span style="color:blue">Explain what values are represented by the size of the bubbles.</span>
<span style="color:black">The larger the bubble, the greater the number of departing flights at that location identified by longitute & latitute.</span>


