---
title: "Homework 1"
author: "CHAN Pok Wah 20259267"
output:
  html_document:
    highlight: tango
    theme: paper
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '3'
---

Note that some of the code blocks contain the expression `eval = FALSE`. This tells R Markdown to display the code contained in the block, but not to evaluate it. Remove it if you want to turn off this behavior.

### Problem 1: Simple logical operations

##### (a) Checking equality.

Given a variable `x`, write a logical expression that evaluates to `TRUE` if the variable `x` is equal to 5369 (the numeric value).  

```{r}
# Insert your code here
x <- 5369
x == 5369
```

##### (b) Checking inequality.

Given a variable `x`, write a logical expression that evaluates to `TRUE` if the variable `x` is *not* `NA` (i.e., is not missing).

```{r}
# Insert your code here
x <- NA
!is.na(x)
```

##### (c) Checking if a number is in a given range. 

Given a (possibly negative) number `x`, write a logical expression that returns `TRUE` if and only if `x` is smaller than `-12` or bigger than `29`.  

```{r}
# Insert your code here
x <- -20
x < -12 || x > 29
```

##### (d) A more complicated example.

Given an integer number `x`, write a logical expression that returns `TRUE` if and only if `x` is an **odd** number between -8 and 12 or 100 and 150.  

```{r }
# Insert your code here
x <- 111
x %% 2 == 1 && ((-8 <= x && x <= 12) || (100 <= x && x <= 150))
```

**Tip**: Recall the modulus operator we saw in lectures: `%%`.  For integers `x` and `y`, `x %% y` is the remainder of `x` divided by `y`.  

### Problem 2: Vector logical operations

##### (a) R has two kinds of logical operators implemented, single (`&`, `|`) and double (`&&`, `||`).  

One of these operators takes advantage of something called *lazy evaluation* while the other does not.  They also don't behave the same way when applied to *vectors*.  

To help you get started, try out the following two examples in your console:

```{r eval = FALSE}
# Example:  The variable y.prob2a is never defined.  
# (Do not define it!)
# What happens when you run this code?
x.prob2a <- 5
(x.prob2a < 10) | (y.prob2a > 2)
(x.prob2a < 10) || (y.prob2a > 2)
```

```{r eval = FALSE}
# Define vectors
x.prob2a.vec <- c(TRUE, FALSE, FALSE)
y.prob2a.vec <- c(FALSE, TRUE, FALSE)

# Apply various logical operations to see what happens
x.prob2a.vec & y.prob2a.vec
x.prob2a.vec && y.prob2a.vec
x.prob2a.vec | y.prob2a.vec
x.prob2a.vec || y.prob2a.vec
```

Please explain what's happening.  Write up a brief explanation below.


<span style="color: darkred; font-size:16pt">
Lazy evaluation is a programming strategy that allows a symbol to be evaluated only when needed. For instance, in the first example (x.prob2a < 10) | (y.prob2a > 2), if x.prob2a contains a value larger or equal to 10, then the second y.prob2a will never be evaluated, thus no exception arised due to Object 'y.prob2a' not defined.
</span>


##### (b) Using `all()` and `any()`

`all()` and `any()` are useful on logical vectors. Since they were not formally introduced in our lectures, you may not know their possible uses. R provides extensive facilities for accessing documentation for R functions, data sets, and other objects, including the `help()` function or `?` help operator. You can use them to access the documentation pages for the two functions and answer the following questions. 

Two people were asked to give their preferences between two options: [Facebook, Twitter], [Firefox, Chrome], [Mac, PC], [Summer, Winter].  Their results are given below.

```{r}
alice.prefs <- c("Twitter", "Chrome", "Mac", "Summer")
bob.prefs <- c("Facebook", "Chrome", "PC", "Summer")
```

Use the `all()` function to determine if the two people have identical preferences. (Your code should ouput a single logical value, either `TRUE` or `FALSE`)

```{r}
# Insert your code here
all(alice.prefs == bob.prefs)
```

Use the `any()` function to determine if the two people have any preferences in common.  (Your code should output a single logical value, either `TRUE` or `FALSE`)

```{r}
# Insert your code here
any(alice.prefs == bob.prefs)
```


##### (c) Missing values.

Let `age` be the vector defined below.

```{r}
age <- c(18, NA, 25, 71, NA, 45, NA, NA, 18)
```

Write a logical expression that checks whether each entry of `age` is missing (recall missing values are denoted by `NA`).  Your expression should return a logical vector having the same length as `age`.

```{r}
# Insert your code here
sapply(age, is.na)
```

### Problem 3: Referencing vector elements

##### (a)  `which()` practice

Again, `which()` was not formally introduced in our lectures. You can use the `help()` function or `?` help operator to access its documentation page and know its usage. Write code that returns the indices of `age` that are missing.

```{r}
# Insert your code here
which(is.na(age))
```

##### (b) Getting non-missing values

Write code that uses negative indices or the negation operator `!` to return only the values of `age` that are *not* missing. (i.e., your code should result in a vector with elements: 18, 25, 71, 45, 18) 


```{r}
# Insert your code here
age[which(!is.na(age))]
```


###Problem 4: Data frame basics

We'll use the `Cars93` dataset in the `MASS` package for this problem. It contains cars selected at random from among 1993 passenger car models.

```{r}
#Remove existing Cars93 to discard previous changes.
rm(Cars93)
library(MASS)
head(Cars93)
str(Cars93)
```

Recall that you can use the `$` operator, the `[` notation, and the `[[` notation to select a column from a data frame. Write code to figure out which cars had an ***MPG.city*** of 35 or more.

```{r}
# Insert your code here
Cars93[Cars93$MPG.city >= 35, ]
```

Use the `which.min()` function and `which.max()` function to figure out the respective car ***Model***s that had the *smallest* and *largest* ***MPG.city***. Your code should return the cars' ***Model***s only. Again, use the `help()` function or `?` help operator to learn the usage of the `which.min()` function and `which.max()` function. 


```{r}
# Insert your code here
Cars93[which.max(Cars93[,'MPG.city']),]
Cars93[which.min(Cars93[,'MPG.city']),]
```


Use the `table()` function to produce a contingency table of observation counts across `DriveTrain` and `AirBags`.

```{r}
# Insert your code here
table(Cars93[,'DriveTrain'], Cars93[,'AirBags'])
```

Use the `levels()` function to code the values contained in the `DriveTrain` column (which is treated as a factor by R). Code `"Rear"` to `"rear wheel"`, `"Front"` to `"front wheel"`, and `"4WD"` to `"all wheel"`. Assign the result back to the `DriveTrain` column.

```{r}
# Insert your code here

#levels(Cars93[,'DriveTrain'])[levels(Cars93[,'DriveTrain'])=="4WD"] <- "all wheel"
#levels(Cars93[,'DriveTrain'])[levels(Cars93[,'DriveTrain'])=="Front"] <- "front wheel"
#levels(Cars93[,'DriveTrain'])[levels(Cars93[,'DriveTrain'])=="Rear"] <- "rear wheel"

levels(Cars93[,'DriveTrain']) <- c("all wheel", "front wheel", "rear wheel") 

```

Use the `ifelse()` function to code the values contained in the `EngineSize` column. Code values greater than 3 to `"large"`, values smaller than 2 than to `"small"`, and remaining values to `"medium"`. Assign the result back to the `EngineSize` column.

```{r}
# Insert your code here
f <- function(x){
  ifelse(x > 3, "large", ifelse(x < 2, "small", "medium"))
}
Cars93$EngineSize <- sapply(Cars93$EngineSize, f)
str(Cars93$EngineSize)
```

The following code defines a list of functions to perform certain transformations on the corresponding columns:

```{r}
trans <- list(
  Price = function(x) x * 0.9,
  EngineSize = function(x) factor(x, labels = c("small", "medium", "large"))
)
```

Then each function can be accessed by using the `[[` notation in the same way as columns of the `Cars93` data frame can be accessed. Complete the loop body below with a **single line** of code to put them into effect on the corresponding columns of `Cars93`:

```{r}
for(var in names(trans)) {
  Cars93[,var] <- trans[[var]](Cars93[,var])
}
```

```{r}
str(Cars93)
```

### Problem 5: A simple summary function

##### (a)  Checking if a vector is binary.

A binary vector is one whose entries are all either 0 or 1.  E.g., `c(0, 0, 1, 1, 0)` is a binary vector, but `c(1, 2, 1.2, 0, 1, 1)` is not.

Write a function called `isBinary` that accepts a single argument called `x`.  Your function should return `TRUE` if `x` is binary and `FALSE` otherwise. You should test your function on at least the example vectors to check that it works as expected.


```{r}
isBinary <- function(x) {
  fe <- function(x) {
    x==1 || x==0
  }
  sum(sapply(x, fe) * rep(TRUE, length=length(x))) == length(x)
}
isBinary(c(0,0,1,0,0))
```


##### (b) What about missing values?

We might also want to treat vectors like `c(0, NA, 1, NA, 0, 1, NA, 1, 1)` as binary.  i.e., we can redefine a binary vector as one whose values are all either 0, 1, or are missing (`NA`).  

Rewrite your `isBinary` function to now have arguments `(x, allow.NA = FALSE)` that has the following behaviour:

1. If `allow.NA` is `TRUE`, your function should return `TRUE` if the elements of `x` are 0, 1 or `NA`.  It should return `FALSE` otherwise.

2. If `allow.NA` is `FALSE` your function should behave exactly as in (a).  i.e., the function should return `TRUE` if and only if the elements of `x` are 0 or 1.  

```{r}
# Insert your code here

isBinary <- function(x, allow.NA = FALSE) {
  if (!allow.NA && any(is.na(x))){
    return(FALSE)
  }
  for (item in x) {
    if (item != 0 && item != 1 && !any(is.na(item))) {
      return(FALSE)
    }
  }
  return(TRUE)
}
isBinary(c(1, 0, 1, NA), TRUE)
isBinary(c(1, 0, 1, NA), FALSE)
isBinary(c(1, 0, 1, 2), FALSE)

```

##### (c) A simple summary function.

Write a function called `calculateBinarySummary` that has a single argument, `x`.  Your function should use your solution to part **(b)** and should have the following behaviour.

1.  If `x` is binary (allowing for `NA`s), your function should return the proportion of values in `x` that are equal to 1, out of the total number of values that are 0 or 1.  (`NA`s should not count toward the denominator in calculating your proportion.)
For example, when `c(NA, 1, 0, 0, 1)` is received, the function should output `0.5`.

2. If `x` is not binary, your function should print a message (using `print()`) telling the user that this function only summarizes binary variables.

```{r}
# Insert your code here
calculateBinarySummary <- function(x) {
  if (isBinary(x, TRUE)) {
    return(mean(x, na.rm = TRUE))
  } else {
    print("This function only summarizes binary variables.")
  }
}
calculateBinarySummary(c(1,1,1,NA))
```

